### 2015-05-01 tex/th1.tex
6 may

\begin{theo} \label{decisionptime}
The decision version of a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$. Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$. Note that $m$ and $k$ are fixed values, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$. Let $\varphi_\A(\X)$ be defined as follows:
\begin{eqnarray*}
\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
\end{eqnarray*}
Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$. Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction. We can do this since checking $\A\models\varphi(\d_i,\X,\e_i)$ can be done in polynomial time for each $i\in\{1,\ldots,n\}$. Note as well that $\varphi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$. Then, let $f'$ be defined as follows:
\begin{eqnarray*}
f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
\end{eqnarray*}
Clearly $f' \in\#\Sigma_0$. Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$. However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time. Therefore, the decision version of $f$ is in {\sc P}.
\end{proof}

>>>>

\begin{theo} \label{decisionptime}
The decision version of a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$. Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$. Note that the values of $m$ and $k$ come from $\varphi(\x,\X,\z)$, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$. Let $\varphi_\A(\X)$ be defined as follows:
\begin{eqnarray*}
\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
\end{eqnarray*}
Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$. Let $\psi_\A(\X)$ be the following formula: For each freely quantified sub-formula $\beta(\x,\z)$ in $\varphi(\x,\X,\z)$ we check if $\A \models \beta(\d_i,\e_j)$, for every $i\in\{1,\ldots,\ell\}$ and $j\in\{1,\ldots,m\}$. If it is true, then we replace $\beta(\d_i,\e_j)$ with a tautology, and otherwise, with a contradiction. Let us give an example on how this would be done: \\

Let $\varphi(\x,\X,\z) = \exists x_1,x_2( X_1(x_2) \wedge ( X_2(z_1,x_1) \wedge \forall y( S_1(y,z_1) )  ))$. In this case, $\x = (x_1,x_2)$, $\X = (X_1,X_2)$ and $\z = (z_1)$. Let $\A\in\Truc$ be $\A = \langle A,S^\A,\leq^\A \rangle$ where $A = \{1,2\}$, $S^\A = \{(1,1),(1,2),(2,2)\}$, and $1 \leq^\A 2$. We generate all the possible assignments for $\x$ and $\z$. And our new formula is:
\begin{multline*}
\varphi_\A(\X) = (X_1(1) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  ) \vee \\
(X_1(2) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  ) \vee \\
(X_1(1) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  ) \vee \\
(X_1(2) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  ).
\end{multline*}
Now we replace the satisfied subformulas:
\begin{multline*}
\psi_\A(\X) = (X_1(1) \wedge X_2(1,1) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,1) \wedge \top  ) \vee \\
(X_1(2) \wedge  X_2(1,1) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,1) \wedge \top  ) \vee \\
(X_1(1) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,2) \wedge \top  ) \vee \\
(X_1(2) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,2) \wedge \top  ),
\end{multline*}
which can be simplified even further:
\begin{multline*}
\psi_\A(\X) = (X_1(1) \wedge X_2(2,1)  ) \vee (X_1(2) \wedge  X_2(2,1) ) \vee \\ (X_1(1) \wedge  X_2(2,2)  ) \vee (X_1(2) \wedge  X_2(2,2)  )
\end{multline*}
Note that we can do this since checking $\A\models\varphi(\d_i,\X,\e_j)$ can be done in polynomial time for each $i,j$. Note as well that $\psi_\A(\X)$ is quantifier-free and its size is also polynomial to the size of $\A$. However, this still is not a valid $\L-$formula since there are no constants in $\L$. Let $P_a$ be a unary predicate that is interpreted as $P_a = \{a\}$ for every $a\in A$. Now let $\zeta_\A(\X,\u)$ be $\psi_\A(\X)$ with the conjunct $P_{a_1}(u_{a_1}) \wedge \cdots \wedge P_{a_{\vert A \vert}}(u_{a_{\vert A \vert}})$, where $u_a$ is a first order variable, and with each of its constants $a\in A$ replaced with $u_a$. In the example:
\begin{multline*}
\zeta_\A(\X,\u) = \big((X_1(u_1) \wedge X_2(u_2,u_1)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_1) ) \vee \\ (X_1(u_1) \wedge  X_2(u_2,u_2)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_2)  )\big) \wedge P_1(u_1) \wedge P_2(u_2),
\end{multline*}
where $\u = (u_1,u_2)$. Now, let $\A^\prime = \A \cup \{P_a\mid a\in A\}$, and let $f'$ be defined as follows:
\begin{eqnarray*}
f'(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \zeta_\A(\P,\e) \} \mid.
\end{eqnarray*}
Clearly $f' \in\#\Sigma_0$. Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A^\prime) > 0$, so computing $f'$ with the instance $\A^\prime$ is enough to solve the decision version of $f$. However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time. Therefore, the decision version of $f$ is in {\sc P}.
\end{proof}

### 2015-05-01 tex/th2.tex
6 may

\begin{theo}
If $\#\Sigma_1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theo}
\begin{proof}
Let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment (satisfying or not) to a 3DNF formula. Suppose that $f_{2^n}-f_{\#3DNF} \in \#\Sigma_1$. This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete. However, as we showed previously in Theorem \ref{decisionptime}, it is also in P. Then, co-NP $\subseteq$ P.
\end{proof}

>>>>

\begin{theo}
If $\#\Sigma_1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theo}
\begin{proof}
Suppose that $\#\Sigma_1$ is closed under substraction, that is, for each pair of functions $f,g\in\#\Sigma_1,$ there is an $h\in\#\Sigma_1$ such that $f(\A)-g(\A) = h(\A)$ for each $\A\in\Truc$.

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $S_i$ is a ternary relation such that $(a_1,a_2,a_3) \in S_i$ iff $(\neg a_1 \wedge \cdots \wedge\neg a_i \wedge a_{i+1} \wedge \cdots \wedge a_3)$ appears as a disjunct in $\psi$, for each $i\in\{1,\ldots,4\}$.

Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (X(x) = X(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. As we supposed initially, let $h\in\#\Sigma_1$ such that $f_{all}-f_{\#3DNF} = h$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ only if $\psi$ is a tautology, so the decision version of $f_{all}-f_{\#3DNF}$ is co-NP-complete. However, as we showed previously in Theorem \ref{decisionptime}, it is also in P. Then, co-NP $\subseteq$ P.
\end{proof}

### 2015-05-01 tex/th3.tex
22 abr
29 abr 1
6 may

\begin{theo}
$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Since $\#\Sigma_1$ is monotonic under addition of elements, $f(\B) \geq 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}

>>>>

\begin{theo}
$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.

He have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\Truc$, if $\A_1$ is an induced substructure of $\A_2$, then $f(\A_1) \leq f(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}. Therefore, $g(\B) \geq g(\A) = 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}
